\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets

### 2018-06-29 doc/paper.tex
final changes to rmd before submission

\if0\blind
{
  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}

  \author{
        Ganesh Krishnan \thanks{The authors gratefully acknowledge funding from CSAFE through
cooperative agreement \# 70NANB15H176 between NIST and Iowa State
University.} \\
    Department of Statistics, Iowa State University\\
     and \\     Heike Hofmann \\
    Department of Statistics and CSAFE, Iowa State University\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
  \end{center}
  \medskip
} \fi

@@

\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
\setlength\parindent{0pt}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{scans-for-land-engraved-areas}{%
\subsection{Scans for land engraved
areas}\label{scans-for-land-engraved-areas}}

@@

\hypertarget{the-chumbley-score-test}{%
\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}

@@

More precisely, let us define starting points \(s_i^{(k)}\) for each
signature \(k = 1, 2\) as \begin{eqnarray}\label{eq.start}
s^{(k)}_i = 
\begin{cases}
t_k^* + i w_v & \text{ for } i < 0 \\
t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
\end{cases}
\end{eqnarray} for integer values of \(i\) with
\(0 < s^{(k)}_i \le T_k - w_v\).

@@

\hypertarget{a-problem-with-failed-tests}{%
\subsection{A problem with failed
tests}\label{a-problem-with-failed-tests}}

@@

\hypertarget{a-modified-approach}{%
\subsection{A modified approach}\label{a-modified-approach}}

@@

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation} i.e.~for an odd number of same-shift correlations, we
skip the middle pair for the different-shift correlations (see
\autoref{sketch-diff-2}). This pairing ensures that the number of
different-shift pairings is the same or at most one less than the number
of same-shift pairings in all tests. In the remainder of the paper, we
will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
and the suggested modified algorithm as \textbf{(CS2)} and compare their
performance on the available scans of the Hamby study.

@@

\hypertarget{testing-setup}{%
\subsection{Testing setup}\label{testing-setup}}

\hypertarget{the-data}{%
\subsubsection{The Data}\label{the-data}}

@@

\hypertarget{setup}{%
\subsubsection{Setup}\label{setup}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{failed-tests}{%
\subsection{Failed Tests}\label{failed-tests}}

@@

\hypertarget{coarseness}{%
\subsection{Coarseness}\label{coarseness}}

@@

\hypertarget{error-rate-assessment}{%
\subsection{Error rate assessment}\label{error-rate-assessment}}

@@

The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized
in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for
false positives (FPR) and false negatives (FNR) are shown side by side
with the area under the curve (AUR) for both methods for a set of
different optimization windows \(w_o\) and a validation window \(w_v\)
of 30 pixels. The rate of false positive same-source identifications is
equal to the statistical type I error rate, which is set to
\(\alpha = 0.05\) for this example. The rate of false negatives are
missed same-source markings. This rate is also known as the type II
error rate.\\
Area under the curve (AUC) is shown with confidence intervals as given
by \citet{delong}. CS2 significantly outperforms CS1 with respect to its
predictive power in most situations.

@@

\hypertarget{observed-versus-nominal-type-i-error-rates}{%
\subsection{Observed versus Nominal Type I error
rates}\label{observed-versus-nominal-type-i-error-rates}}

@@

\hypertarget{high-resolution-hamby-44-scans}{%
\subsection{High resolution Hamby 44
scans}\label{high-resolution-hamby-44-scans}}

@@

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

@@

\section{Scenarios of failed tests}

Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned. These failures are inherent to the Chumbley Score and cannot be prevented.
\begin{figure}[hbtp]
\centering
\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}

@@

\newpage
\section{Type 2 errors: CS1 vs CS2}

\autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.

>>>>

\if0\blind
{
  \title{\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
bullets}

  \author{
        Ganesh Krishnan,\\ %\\
    Master of Science (M.Sc.)\\
    % \\
     and \\     Prof.~Heike Hofmann,\\ %\\
    Doctor of Philosophy (Ph.D.)\\
    % \\
      }
    \maketitle
  \bigskip
  \begin{center}
  {\textbf{Affiliation(s)}\\
  Department of Statistics, Iowa State University and CSAFE (Center for
  Statistics and Applications in Forensic Evidence)\\
  \bigskip
  \textbf{Source of funding}\\
  The authors gratefully acknowledge funding from CSAFE through
  cooperative agreement \# 70NANB15H176 between NIST and Iowa State
  University.\\
  \bigskip
  \textbf{Acknowledgement}\\
  We would like to thank the efforts of CSAFE and Roy J Carver Microscopy
  lab in scanning the Hamby-44 set and providing the scans to us. We would
  also like to thank NIST for providing the Hamby-44 bullets to us for
  scanning.\\
  \bigskip
  \textbf{Information on presentation of work}\\
  This work was presented in the form of a poster in the AFTE 2018 49th
  Annual Training Seminar, held from June 3rd to June 8th, 2018 in
  Charleston, West Virginia}
  \end{center}

} \fi

  % \bigskip

  % \bigskip
\newpage

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
  \title{\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
bullets}
  \end{center}
  \bigskip
  \maketitle
} \fi

% \if1\blind
% {
%   \bigskip
%   \bigskip
%   \bigskip
%   \begin{center}
%     {\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
bullets}
%   \end{center}
%   \medskip
% } \fi

@@

\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
\newcommand{\cited}[1]{{\textcolor{red}{#1}}}

\setlength\parindent{0pt}

@@

\section{Introduction}\label{introduction}

@@

\section{Methods}\label{methods}

\subsection{Scans for land engraved
areas}\label{scans-for-land-engraved-areas}

@@

\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}

@@

More precisely, let us define starting points \(s_i^{(k)}\) for each
signature \(k = 1, 2\) as

\begin{eqnarray}\label{eq.start}
s^{(k)}_i = 
\begin{cases}
t_k^* + i w_v & \text{ for } i < 0 \\
t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
\end{cases}
\end{eqnarray}

for integer values of \(i\) with \(0 < s^{(k)}_i \le T_k - w_v\).

@@

\subsection{A problem with failed
tests}\label{a-problem-with-failed-tests}

@@

\subsection{A modified approach}\label{a-modified-approach}

@@

We then define the pairs for different-shifts as

\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}

i.e.~for an odd number of same-shift correlations, we skip the middle
pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the
same or at most one less than the number of same-shift pairings in all
tests. In the remainder of the paper, we will refer to the algorithm
defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified
algorithm as \textbf{(CS2)} and compare their performance on the
available scans of the Hamby study. The section
\ref{appendix:appxfailed} of the Appendix discusses about the various
scenarious of failed tests.

@@

\subsection{Testing setup}\label{testing-setup}

\subsubsection{The Data}\label{the-data}

@@

\subsubsection{Setup}\label{setup}

@@

\section{Results}\label{results}

\subsection{Failed Tests}\label{failed-tests}

@@

\subsection{Coarseness}\label{coarseness}

@@

\subsection{Error rate assessment}\label{error-rate-assessment}

@@

The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized
in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for
false positives (FPR) and false negatives (FNR) are shown side by side
with the area under the curve (AUR) for both methods for a set of
different optimization windows \(w_o\) and a validation window \(w_v\)
of 30 pixels. The rate of false positive same-source identifications is
equal to the statistical type I error rate, which is set to
\(\alpha = 0.05\) for this example. The rate of false negatives are
missed same-source markings. This rate is also known as the type II
error rate. A detailed plot on the type II error rates for CS1 and CS2
can be found in the section \ref{appendix:appxtype2} of the Appendix.
Area under the curve (AUC) is shown with confidence intervals as given
by \citet{delong}. CS2 significantly outperforms CS1 with respect to its
predictive power in most situations.

@@

\subsection{Observed versus Nominal Type I error
rates}\label{observed-versus-nominal-type-i-error-rates}

@@

\subsection{High resolution Hamby 44
scans}\label{high-resolution-hamby-44-scans}

@@

\section{Conclusions}\label{conclusions}

@@

\section{Appendix}\label{appendix}

@@

\subsection{Scenarios of failed tests}
\label{appendix:appxfailed}
Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned. These failures are inherent to the Chumbley Score and cannot be prevented.
\begin{figure}[hbtp]
\centering
\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}

@@

\newpage
\subsection{Type 2 errors: CS1 vs CS2}
\label{appendix:appxtype2}
\autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
